---
title: Useful R code — `RMarkdown`  
author: |
 | Matthew Malishev
 | @darwinanddavis 
fontsize: 10
geometry: margin=1in
documentclass: article
linkcolor: pink
urlcolor: blue
citecolor: red
always_allow_html: yes
output:
  html_document:
    highlight: tango
    code_folding: show
    toc: yes
    toc_depth: 4
    number_sections: no
    toc_float: yes
  md_document: 
    variant: markdown_github
  pdf_document:
    includes:
      in_header: # add .tex file with header content
    highlight: tango
    template: null
    toc: yes
    toc_depth: 4
    number_sections: false
    fig_width: 4
    fig_height: 5
    fig_caption: true
    df_print: tibble 
    citation_package: biblatex # natbib
    latex_engine: xelatex #pdflatex # lualatex
    keep_tex: true # keep .tex file in dir 
  word_document:
    highlight: tango
    keep_md: yes
    pandoc_args: --smart
    #reference: mystyles.docx
    toc: yes
    toc_depth: 4
inludes:
  before_body: before_body.tex
subtitle: 
tags:
- nothing
- nothingness
params: 
  dir: "/Users/malishev/Documents/Melbourne Uni/Programs/R code/UsefulCode"
  date: !r Sys.Date()
  version: !r getRversion()
  email: "matthew.malishev@gmail.com"
  doi: https://github.com/darwinanddavis/UsefulCode
classoption: portrait
# ^['https://github.com/darwinanddavis/UsefulCode'] # footnote
vignette: >
  %\VignetteIndexEntry{Useful R code}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

<script type="text/x-mathjax-config">
  MathJax.Hub.Config({ TeX: { equationNumbers: {autoNumber: "all"} } });
</script>

```{r echo = FALSE}
 library(rmarkdown)
# setwd("")
# f <- list.files()[1]
# render(f, output_format='pdf_document')
# render(f, output_format='pdf_document')
```

```{r, set-options, echo = F, cache = FALSE}
options(width=100)
knitr::opts_chunk$set(
 eval = F, # run all code
 results='hide',
 echo = T, # show code chunks in output
 comment = "",
 tidy.opts=list(width.cutoff=100), # set width of code chunks in output
 tidy=TRUE, # make output as tidy
 message = FALSE,  # mask all messages
 warning = FALSE, # mask all warnings 
 size="small" # set code chunk size
)

# https://github.com/ucb-stat133/stat133-fall-2016/blob/master/hws/hw02-tables-ggplot.Rmd
knitr::opts_knit$set(root.dir=paste0(params$dir,"/")) # set working dir

setwd(paste0(params$dir,"/")) # for running just in R not knitr
```

\  

Date: `r params$date`  
`R` version: `r params$version`  
*Corresponding author: `r params$email`  
This document can be found at `r params$doi`

\newpage  

## Overview

Useful functions and code snippets for building `RMarkdown` docs. _All `R` code is in Rmd file._       


<!-- ———————————————————————————————————————————————————————————————————————————— -->
***  
### Split page into three columns (displays best in browser).  
_`R` code is in Rmd file._    

<!-- #https://holtzy.github.io/Pimp-my-rmd/ -->
<div class = "row">

<div class = "col-md-4">
```{r, rmarkdown1, message=F, echo=T, eval=T}
# r plot code
require(ggplot2)
ggplot(mtcars, aes(x=mpg)) + 
  geom_histogram(fill="skyblue", alpha=0.5) + 
  theme_classic()
```
</div>
  
<div class = "col-md-4">
<br><br>  

>Praise the lord, I was born to travel  
>Feeling like Slash in front of the chapel  
>I'm leaned back with the Les Paul  
>Shit I smoke is like cholesterol  
>Spilled dressin' on the vest at the festival  
>The best of all, had a midget Puerto Rican at my beckon call

</div>
  
<div class = "col-md-4">
<br><br>
Pump the bass in the trunk  
It rattled like a baby hand  
Except this toy cost 80 grand  
And I'm crazy tan, from all the places that I've been  
Just from writing words with a pen  

</div>
<!-- close div -->
</div> 

<!-- ———————————————————————————————————————————————————————————————————————————— -->
\  
\  

### Inserting text within indented equations    
$$
p(x) = \theta^{x} (1 - \theta)^{1-x} ~~\mbox{ for }~~x = 0,1  
$$

### Remove white space around imgs using `fig.asp`
`{r,fig.asp=0.50}`  

### Read external Rmd chunks   
<!-- https://stackoverflow.com/questions/52397430/include-code-from-an-external-r-script-run-in-display-both-code-and-output -->

**Option 1**  
Add this to code chunk: `{r, code = readLines("file.R")}`  

**Option 2**
<!-- live example in 'olympics.Rmd' -->
```{r, eval = F, message=F}
# load r datasets and code chunks 
require(dplyr)
paste0(params$here) %>% 
  list.files("r",full.names = T) %>%
  sapply(knitr::read_chunk)

# or 
knitr::read_chunk('external.R')
```

Then in 'external.R' file, add the following header line:   
`# @knitr external`    
Then use add this line to the code chunk in the main Rmd file:   
`<<external>>`  


### Knit an .R file script    
<!-- https://twitter.com/jamespwr/status/1389916296171724803 -->
