---
title: Useful R code - sf and geodata  
author: |
 | Matthew Malishev^1^*  
fontsize: 10
geometry: margin=1in
documentclass: article
linkcolor: pink
urlcolor: blue
citecolor: red
always_allow_html: yes
output:
  html_document:
    highlight: tango
    code_folding: show
    toc: yes
    toc_depth: 4
    number_sections: no
    toc_float: yes
  md_document: 
    variant: markdown_github
  pdf_document:
    includes:
      in_header: # add .tex file with header content
    highlight: tango
    template: null
    toc: yes
    toc_depth: 4
    number_sections: false
    fig_width: 4
    fig_height: 5
    fig_caption: true
    df_print: tibble 
    citation_package: biblatex # natbib
    latex_engine: xelatex #pdflatex # lualatex
    keep_tex: true # keep .tex file in dir 
  word_document:
    highlight: tango
    keep_md: yes
    pandoc_args: --smart
    #reference: mystyles.docx
    toc: yes
    toc_depth: 4
inludes:
  before_body: before_body.tex
subtitle: 
tags:
- nothing
- nothingness
params: 
  dir: "/Users/malishev/Documents/Melbourne Uni/Programs/R code/UsefulCode"
  date: !r Sys.Date()
  version: !r getRversion()
  email: "matthew.malishev [at] gmail.com"
  doi: https://github.com/darwinanddavis/UsefulCode
classoption: portrait
# ^['https://github.com/darwinanddavis/UsefulCode'] # footnote
vignette: >
  %\VignetteIndexEntry{Useful R code}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

<script type="text/x-mathjax-config">
  MathJax.Hub.Config({ TeX: { equationNumbers: {autoNumber: "all"} } });
</script>

```{r echo = FALSE}
 library(rmarkdown)
# setwd("")
# f <- list.files()[1]
# render(f, output_format='pdf_document')
# render(f, output_format='pdf_document')
```

```{r, set-options, echo = FALSE, cache = FALSE}
options(width=100)
knitr::opts_chunk$set(
 eval = F, # run all code
 results='hide',
 # echo = FALSE, # show code chunks in output 
 comment = "",
 tidy.opts=list(width.cutoff=100), # set width of code chunks in output
 tidy=TRUE, # make output as tidy
 message = FALSE,  # mask all messages
 warning = FALSE, # mask all warnings 
 collapse = T,
 size="small" # set code chunk size
)

# https://github.com/ucb-stat133/stat133-fall-2016/blob/master/hws/hw02-tables-ggplot.Rmd
knitr::opts_knit$set(root.dir=paste0(params$dir,"/")) # set working dir

setwd(paste0(params$dir,"/")) # for running just in R not knitr
pacman::p_load(dplyr,readr,rvest,xml2,magrittr,sp,sf,rgdal,ggmap,ggplot2,stringr,ggthemes,ggnetwork,colorspace,ggtext,ggsn,ggspatial,showtext,here)
```

\  

Date: `r params$date`  
`R` version: `r params$version`  
*Corresponding author: `r params$email`  
This document can be found at `r params$doi`

\newpage  

## Overview

Accessing sf, sp, geos,a nd geojson objects, such as attributes, accessing lower levels of their structure, latlon and MULTIPOLYGONS layers, etc.  

<!-- ideas -->
<!-- https://r-spatial.github.io/sf/articles/sf1.html -->
<!-- https://nceas.github.io/oss-lessons/spatial-data-gis-law/3-mon-intro-gis-in-r.html -->

## sf simple features  
Get bbox coords      
```{r,collapse=T}
require(sf)
methods(class="sf")
obj$geometry %>% attr(which = "bbox")

```

Get first element or range of sf  
```{r}
obj$geometry
obj$geometry[1] %>% unlist %>% .[1] # first element
obj %>% st_coordinates() %>% .[1:3] # coord range    
```

Get polygons (and range) in geometry layer       
```{r}
obj %>% st_geometry() %>% .[1:10] # polygon range  
```


### .shp files    

Read in .shp file    
```{r, collapse=T}
require(rgdal)# \ rgdal 
shp <- readOGR(".shp", layer="X") # layer = name of geometry element
```

Get and transform projection  
```{r}
require(rgdal)# \ rgdal 
proj4string(shp) # get projection 
spTransform(shp, CRS("+proj=longlat +datum=WGS84")) # transform projection 
```

### Base maps       
Base map from OSM    
```{r}
# osm 
gg <- get_map(location,
              source="osm", 
              color="bw")

ggmap(gg, extent = "device", darken = c(0.9,"white")) +
  geom_point(data = df, aes(x,y))
```



