---
title: Useful R code - sf and geodata  
author: |
 | Matthew Malishev^1^*  
fontsize: 10
geometry: margin=1in
documentclass: article
linkcolor: pink
urlcolor: blue
citecolor: red
always_allow_html: yes
output:
  html_document:
    highlight: tango
    code_folding: show
    toc: yes
    toc_depth: 4
    number_sections: no
    toc_float: yes
  md_document: 
    variant: markdown_github
  pdf_document:
    includes:
      in_header: # add .tex file with header content
    highlight: tango
    template: null
    toc: yes
    toc_depth: 4
    number_sections: false
    fig_width: 4
    fig_height: 5
    fig_caption: true
    df_print: tibble 
    citation_package: biblatex # natbib
    latex_engine: xelatex #pdflatex # lualatex
    keep_tex: true # keep .tex file in dir 
  word_document:
    highlight: tango
    keep_md: yes
    pandoc_args: --smart
    #reference: mystyles.docx
    toc: yes
    toc_depth: 4
inludes:
  before_body: before_body.tex
subtitle: 
tags:
- nothing
- nothingness
params: 
  dir: "/Users/malishev/Documents/Melbourne Uni/Programs/R code/UsefulCode"
  date: !r Sys.Date()
  version: !r getRversion()
  email: "matthew.malishev [at] gmail.com"
  doi: https://github.com/darwinanddavis/UsefulCode
classoption: portrait
# ^['https://github.com/darwinanddavis/UsefulCode'] # footnote
vignette: >
  %\VignetteIndexEntry{Useful R code}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

<script type="text/x-mathjax-config">
  MathJax.Hub.Config({ TeX: { equationNumbers: {autoNumber: "all"} } });
</script>

```{r echo = FALSE}
 library(rmarkdown)
# setwd("")
# f <- list.files()[1]
# render(f, output_format='pdf_document')
# render(f, output_format='pdf_document')
```

```{r, set-options, echo = FALSE, cache = FALSE}
options(width=100)
knitr::opts_chunk$set(
 eval = F, # run all code
 results='hide',
 # echo = FALSE, # show code chunks in output 
 comment = "",
 tidy.opts=list(width.cutoff=100), # set width of code chunks in output
 tidy=TRUE, # make output as tidy
 message = FALSE,  # mask all messages
 warning = FALSE, # mask all warnings 
 collapse = T,
 size="small" # set code chunk size
)

# https://github.com/ucb-stat133/stat133-fall-2016/blob/master/hws/hw02-tables-ggplot.Rmd
knitr::opts_knit$set(root.dir=paste0(params$dir,"/")) # set working dir

setwd(paste0(params$dir,"/")) # for running just in R not knitr
pacman::p_load(dplyr,readr,rvest,xml2,magrittr,sp,sf,rgdal,ggmap,ggplot2,stringr,ggthemes,ggnetwork,colorspace,ggtext,ggsn,ggspatial,showtext,here)
```

\  

Date: `r params$date`  
`R` version: `r params$version`  
*Corresponding author: `r params$email`  
This document can be found at `r params$doi`

\newpage  

## Overview

Accessing sf, sp, geos,a nd geojson objects, such as attributes, accessing lower levels of their structure, latlon and MULTIPOLYGONS layers, etc.  

<!-- ideas -->
<!-- https://r-spatial.github.io/sf/articles/sf1.html -->
<!-- https://nceas.github.io/oss-lessons/spatial-data-gis-law/3-mon-intro-gis-in-r.html -->

### sf simple features  
Get bbox coords      
```{r,collapse=T}
require(sf)
methods(class="sf")
obj$geometry %>% attr(which = "bbox")

```

Get first element or range of sf  
```{r}
obj$geometry
obj$geometry[1] %>% unlist %>% .[1] # first element
obj %>% st_coordinates() %>% .[1:3] # coord range    
```

Get polygons (and range) in geometry layer       
```{r}
obj %>% st_geometry() %>% .[1:10] # polygon range  
```


### .shp files    

Read in .shp file    
```{r, collapse=T}
require(rgdal)# \ rgdal 
shp <- readOGR(".shp", layer="X") # layer = name of geometry element
```

Get and transform projection  
```{r}
require(rgdal)# \ rgdal 
proj4string(shp) # get projection 
spTransform(shp, CRS("+proj=longlat +datum=WGS84")) # transform projection 
```

### Base maps       
Base map from OSM    
```{r}
# osm 
gg <- get_map(location,
              source="osm", 
              color="bw")

ggmap(gg, extent = "device", darken = c(0.9,"white")) +
  geom_point(data = df, aes(x,y))
```

### North arrow and scale bars  
<!-- https://www.r-spatial.org/r/2018/10/25/ggplot2-sf.html   -->
```{r}
library("ggspatial")
ggplot(data = world) +
    geom_sf() +
    annotation_scale(location = "bl", width_hint = 0.5) +
    annotation_north_arrow(location = "bl", which_north = "true", 
        pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"),
        style = north_arrow_fancy_orienteering) +
    coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.97))

## Scale on map varies by more than 10%, scale bar may be inaccurate
```



```{r}

# accessing coords in sf (complex)
# shp <- here::here("worldmaps","data","day14.shp") %>% readOGR()
# d <- shp@data # df vars
# bb <- shp@bbox #bbox
# ply <- shp@polygons # polygons 
# plyl <- ply %>% purrr::map(ggplot2::fortify)  # melt polygon class
# names(plyl) <- d$VARNAME_3 # get district name  
# ggplot() + 
#   geom_polygon(data = ply_df[[600]],
#                aes(long, lat)) 

# ply[[1]]@Polygons[[1]]@coords # works
# ply[[1]] %>% attr(which="Polygons") %>% purrr::map("coords") # works

```

<!-- ### Changing fonts in ggplot -->
<!-- https://stackoverflow.com/questions/34522732/changing-fonts-in-ggplot2 -->


### Projections  

```{r}

#list of proj 
# https://proj.org/operations/projections/patterson.html
# https://spatialreference.org/ref/?search=albers&srtext=Search


require(mapdata,rnaturalworld)
### 
d <- ne_countries(scale = "large",
             type = "countries",
             country = "Japan",
             returnclass = "sf")
bb <- d$geometry %>% st_bbox()

# get current proj 
d %>% st_crs() # /sf
d %>% crs() # /raster

# transform proj options 
crsn <- 3995 # 3995 3033
plat <- 50
plon <- 0
prj <- paste0("+lat_0=",plat," +lon_0=",plon," +init=epsg:",crsn) # opt1
prj <- "+init=epsg:4326" # opt2

# transforming for diff data types
ras@crs@projargs <- prj # raster
dl %>% st_transform(prj) # shp


dt <- d %>% st_transform("+proj=utm +zone=19 +ellps=GRS80 +datum=NAD83")
dt <- d %>% st_transform(crsn)

# quartz()
p <- ggplot(data = d) +
  geom_sf() +
  coord_sf(crs = prj) # option 1 
  coord_sf(crs = st_crs(crsn)) # option 2



```

### KML/KMZ data  
<!-- https://mitchellgritts.com/posts/load-kml-and-kmz-files-into-r/ -->
```{r}
require(sf)
path <- "/data.kml" %>% sf::st_read() # single layer kml
path <- "/data.kml" %>% sf::st_layers() # multi layer kml

```


### Convert sp/sf/spdf data to data frame (fortify)  
```{r}
require(rworldmap)
require(broom)
d <- getMap(resolution = "high") # get world data
d_fort <- tidy(d, region = "NAME") # fortify 
```

### Find intersecting geo points within sf objects
```{r}
require(osmdata)
require(sf)
get_osm_feature <- function(loc,key,value,geo){
  oo <- opq(loc) %>% # find location
    add_osm_feature(key = key, value = value) %>% # source key-value pair
    osmdata_sf() %>% # get sf obj
    unname_osmdata_sf() # combat sf error
  oo[[geo]] # %>% # pull geometry
}

# location, key, value 
location1 <- "melbourne australia" 
polyg <- get_osm_feature(location1, "leisure","park|natu","osm_polygons") # get osm features

# check if point intersects with polygon 
pts_inter = c(144.7695,-38.2724) %>% 
  st_point() %>% # convert into sf point
  st_sfc(crs = 4326) %>% # set prj
  st_as_sf() # convert into sf  
polyg  <- polyg %>% st_transform(4326) # match proj to above sf coords
polyg_inter <- st_intersection(polyg,pts_inter) # find intersecting point within polygon 

# plot intersecting polygon and point
st_intersection(pts_inter,polyg) %>% st_geometry()
polyg %>% filter(osm_id == "234385754") 
ggplot() +
  geom_sf(data = polyg %>% filter(osm_id == polyg_inter$osm_id)) + # plot polygon w intersecting point
  geom_sf(data = pts_inter, col = "red") + # add intersecting point
  geom_sf_text(data = pts_inter, label = "Intersecting\npoint") +
  theme_bw()

```

Other predicate funcs for sf pairs
```{r, eval = F}
st_intersects()
st_disjoint()
st_touches()
st_crosses()
st_within()
st_contains()
st_contains_properly()
st_overlaps()
st_equals()
st_covers()
st_covered_by()
st_equals_exact()
st_is_within_distance()
```

### Convert df/numeric/matrix to geometry

Convert points to linestring
```{r}
inter_df <- rbind(df1,df2,df3) %>%   
    as.matrix() %>%
    st_linestring(dim = "XY") %>% # convert points to single linestring
    st_sfc(crs = 4326) # convert sfg object to geometry 
```

Convert each row in df to linestring
```{r}
inter_df <- rbind(df1,df2,df3) %>%
    st_as_sf(coords = c("lon","lat"), crs = 4326) %>% 
    st_geometry() %>% 
    st_cast("LINESTRING")
```

