---
title: Useful R code in `ggplot`
author: |
 | Matthew Malishev^1^*
 |  
 | _^1^ Department of Biology, Emory University, 1510 Clifton Road NE, Atlanta, GA, USA, 30322_
#bibliography:/Users/malishev/Documents/Melbourne Uni/Thesis_2016/library.bib
fontsize: 10
geometry: margin=1in
documentclass: article
linkcolor: pink
urlcolor: blue
citecolor: red
always_allow_html: yes
output:
  html_document:
    highlight: tango
    code_folding: show
    toc: yes
    toc_depth: 4
    number_sections: no
    toc_float: yes
  md_document: 
    variant: markdown_github
  pdf_document:
    includes:
      in_header: # add .tex file with header content
    highlight: tango
    template: null
    toc: yes
    toc_depth: 4
    number_sections: false
    fig_width: 4
    fig_height: 5
    fig_caption: true
    df_print: tibble 
    citation_package: biblatex # natbib
    latex_engine: xelatex #pdflatex # lualatex
    keep_tex: true # keep .tex file in dir 
  word_document:
    highlight: tango
    keep_md: yes
    pandoc_args: --smart
    #reference: mystyles.docx
    toc: yes
    toc_depth: 4
inludes:
  before_body: before_body.tex
subtitle: 
tags:
- nothing
- nothingness
params: 
  dir: "/Users/malishev/Documents/Melbourne Uni/Programs/R code/UsefulCode"
  date: !r Sys.Date()
  version: !r getRversion()
  email: "matthew.malishev@gmail.com"
  doi: https://github.com/darwinanddavis/UsefulCode
classoption: portrait
# ^['https://github.com/darwinanddavis/UsefulCode'] # footnote
vignette: >
  %\VignetteIndexEntry{Useful R code}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

<script type="text/x-mathjax-config">
  MathJax.Hub.Config({ TeX: { equationNumbers: {autoNumber: "all"} } });
</script>

```{r echo = FALSE}
 library(rmarkdown)
# setwd("")
# f <- list.files()[1]
# render(f, output_format='pdf_document')
# render(f, output_format='pdf_document')
```

```{r, set-options, echo = FALSE, cache = FALSE}
options(width=100)
knitr::opts_chunk$set(
 eval = F, # run all code
 results='hide',
 # echo = FALSE, # show code chunks in output 
 comment = "",
 tidy.opts=list(width.cutoff=100), # set width of code chunks in output
 tidy=TRUE, # make output as tidy
 message = FALSE,  # mask all messages
 warning = FALSE, # mask all warnings 
 size="small" # set code chunk size
)

# https://github.com/ucb-stat133/stat133-fall-2016/blob/master/hws/hw02-tables-ggplot.Rmd
knitr::opts_knit$set(root.dir=paste0(params$dir,"/")) # set working dir

setwd(paste0(params$dir,"/")) # for running just in R not knitr
```

\  

Date: `r params$date`  
`R` version: `r params$version`  
*Corresponding author: `r params$email`  
This document can be found at `r params$doi`

\newpage  

## Overview

Same deal as Useful Code 1 and 2 except just `gglot` because it's too difficult sifting through the other docs.  

## `ggplot` functions  
### Remove annoying stock gridlines from plot window  
```{r, ggplot1, results='hide',eval=F}
plot + theme_bw() + 
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                            panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
# alternative (after loading ggridges library)
theme_ridges(grid=F,center_axis_labels = T)
```

### Setting global graphics theme for ggplot
```{r, ggplot2, results='hide',eval=F}
plot_it_gg <- function(bg,family){ # bg = colour to plot bg, family = font family
  theme_tufte(base_family = family) +
  theme(panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = bg,
                                        colour = bg),
        plot.background = element_rect(fill=bg)
  ) +
    theme(axis.line = element_line(color = "white")) +
    theme(axis.ticks = element_line(color = "white")) +
    theme(plot.title = element_text(colour = "white")) +
    theme(axis.title.x = element_text(colour = "white"), 
          axis.title.y = element_text(colour = "white")) +
    theme(axis.text.x = element_text(color = "white"),
          axis.text.y = element_text(color = "white")) +
    theme(legend.key = element_rect(fill = bg)) + 
    theme(legend.title = element_text(colour="white")) + 
    theme(legend.text = element_text(colour="white")) 
} 
```

### Put plot in function to take dynamic data inputs  
Ref: http://jcborras.net/carpet/visualizing-political-divergences-2012-local-elections-in-helsinki.html
```{r, ggplot3, results='hide',eval=F}
hr.mass.plot <- function(d) {
  p <- ggplot(d, aes(HR, Mass, color = colfunc)) + 
    geom_density_2d(data=d, aes(x = HR, y = Mass), 
                    stat = "density2d",position="identity", 
                    color=adjustcolor("orange",alpha=0.8), size=1.5, contour = T, lineend="square",linejoin="round") 
  p <- p + geom_point(data=d, aes(x = HR, y = Mass),
                      color=colfunc,
                      fill=colfunc) +
    scale_color_manual(values = magma(8))
  p <- p + scale_y_continuous(limits=c(-200,200), name="Mass lost (g)") 
  p <- p + scale_x_continuous(limits=c(0,0.35),name=expression("Home range area (km^2)")) 
  p <- p + theme_classic()
  print(p)
}
hr.mass.plot(d)
```

### Using `ggplot` when looping through `for` loop and saving to dir  
```{r, ggplot4,eval=F}
pdf("mypdf.pdf",onefile = T)
for(i in 1:3){ 
par(bty="n", las = 1)
  grid.arrange( 
  ggplot(data, aes(x = X, y = Y, fill=..x..)) + 
    geom_density_ridges_gradient(scale = 5, size=0.2,color="black", rel_min_height = 0.01,panel_scaling=T,alpha=0.2) +
    geom_density_ridges(scale = 5, size=0.2,color="black", rel_min_height = 0.01,fill="white",alpha=0.2) +
    scale_fill_viridis(name = "Diameter", alpha=0.1, option = "magma",direction=-1) + 
    xlim(c(0,25)) +
    labs(title = paste0("Title_",i)) +
    xlab("X") +
    ylab("Y")
  )
} 
# end loop 
dev.off()


# geom_density_ridges() # scale = overlap

# geom_density_ridges(scale = 5, size=0.2,color="white", rel_min_height = 0.01,fill=col,alpha=0.5) +

# scale_fill_viridis option =  "magma", "inferno","plasma", "viridis", "cividis"

```

### Converting lists and dataframes to usable format for `ggplot` (`melt` package)
```{r, ggplot5, results='hide',eval=F}
# ------------------------- plot individual outputs -------------------------
mm_ = readRDS(paste0(model.path,fh,".R"))
cat("order = cerc, food, juv, adult, infected, infected shedding, host length, parasite mass")
# plot master
mm <- mm_[[2]]
y_m <- melt(mm);y_m
ggplot() +
  geom_point(data = y_m, aes(x = rep.int(1:n.ticks,max(L1)) , y = value, group = L1, colour=factor(L1)), ) +
  geom_line(data = y_m, aes(x = rep.int(1:n.ticks,max(L1)) , y = value, group = L1, colour=factor(L1)), ) +
  #linetype=y_m$L1) +
  theme_tufte() 
# +  geom_text(x=,y=,label = max(value),check_overlap = TUE)

```

### Insert math expression in legend title      
```{r, ggplot6, results='hide',eval=F}
ggplot() + 
  labs(title = bquote("Hello" ~ r[xy] ~ "and" ~ B^2))
```

### Create double line break with expression in legend title (and labels)      
```{r, ggplot7, results='hide',eval=F}
ggplot() +
  scale_color_manual(
    expression(atop("text", 
    atop(textstyle(epsilon))))
    )


ggplot() +
  scale_color_manual(
    name=expression(
      atop("Productivity",
           atop(textstyle(
          (mg~C~L^{-1}~d^{-1}) # this is bracketed text
          ))))
  )

```

### Adding text lables to plots
```{r, gg8,eval=T}
require(ggplot2)
xx = sample(100, 100)
yy = rnorm(100)

ggplot() +
  geom_point(aes(xx,yy,size=5),colour=xx,show.legend = F) +
  geom_text(aes(xx,yy,label=xx),check_overlap = T,size=yy+5) +
  theme_classic()

# add bg label
ggplot() +
  geom_line(aes(xx,yy,size=5),colour=xx,show.legend = F) +
  geom_label(aes(xx,yy,label=xx),size=yy+5,color = yy +5, fill=xx) +
  theme_classic()
```

### Calling data frame columns with weird formatting 
```{r,gg9}
# use ticks, not apostrophes  
ggplot(df, aes(X, `Y col with spaces`)) + 
  geom_line()
```

### Use POSIX format
```{r,gg10}
# turn into posix month day year format
# Date = factor
require(lubridate)
df$Date<- mdy_hms(df$Date, tz='GMT')

ggplot(df, aes(Date,Y)) + geom_line() +
  scale_x_datetime(date_breaks = "1 month", 
                   date_minor_breaks = "1 week", 
                   date_labels = "%B %Y")
```

### Plotting multiple plots per window (with different plot size ratios)  
```{r,gg11}
# 3 plots above, 1 below
grid.arrange(
  grobs = ggplot_list, # list with ggplots or grobs
  widths = c(1, 1, 1),
  layout_matrix = rbind(c(1, 2, 3),
                        c(4, 4, 4))
)

# 3 plots above with first plot 2 plots wide 
# 2 plots below with third plot 3 plots wide 
grid.arrange(
  grobs = ggplot_list, # list with ggplots or grobs
  widths = c(2, 1, 1),
  layout_matrix = rbind(c(1, 2, NA),
                        c(3, 3, 4))
)
```

Setting legend to custom aes elements, e.g. color, shape, linetype
```{r,gg12,eval=T}
# set 'labs' arguments to same title as in 'scale_color_manual'
# legend values automatically matches arguments passed to colvec
# NB specifying the labels arg in scale_color_manual overrides labs  

require(ggplot2)

snack_df <- data.frame(
  "X"=sample(100,10,replace=F),
  "Y"=sample(100,10,replace=F),
  "Sum"=runif(10),
  "Size"=rep(LETTERS[1:5],each=2)
)
colvec <- colorspace::sequential_hcl(length(unique(snack_df$Size)), "SunsetDark")
legend_ttl <- "This is your legend"
legend_pars <- unique(snack_df$Size) # not run
                       
ggplot(snack_df,aes(X,Y)) +
  geom_line(aes(group=Size,color=Size,linetype=Size),size=1) +
  geom_point(aes(group=Size,color=Size,shape=Size),size=3) +
  scale_color_manual(name=legend_ttl,
                     # labels = legend_pars, # this overrides labs 
                     values = colvec) +
  geom_text(aes(X,Y,
                label=c(Size)
                ),
            check_overlap = T,
            size=5,vjust=-1,hjust=1.2) +
  labs(title="Plot title",x="X",y="Y",
       # ----- these are the legend key arguments
       colour=legend_ttl,
       fill=legend_ttl,
       linetype=legend_ttl,
       shape=legend_ttl) + # check shape 
  theme_classic() 

```

